---
title: "Question 2"
author: "Romulo Jimenez"
date: "2025-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Hypothesis Testing – HHP Crosstabs"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

## Question 2

### Crosstab 

| Education | Gender | n_12 | n_34 |
|:---------:|:------:|-----:|-----:|
| no_coldeg | male   | 13534 | 4425 |
| no_coldeg | female | 20840 | 7569 |
| no_coldeg | tran/o | 322 | 252 |
| coldeg    | male   | 40663 | 6757 |
| coldeg    | female | 60524 | 11448 |
| coldeg    | tran/o | 757 | 452 |

### Hypothesis test (college degree: female vs. male)

We test whether the proportion of higher DOWN responses (3–4) differs by gender among college-degree holders.  
Null: \(H_0: p_{\text{male}} = p_{\text{female}}\).  
Alt: \(H_a: p_{\text{male}} \neq p_{\text{female}}\).

Among college-degree holders, 6,757 of 47,420 males and 11,448 of 71,972 females reported 3–4 (14.25% vs. 15.90%).  
The estimated difference is 0.0165. Using a pooled standard error of about 0.0027 gives \(z \approx 6.1\) and \(p < 0.001\).  
We reject \(H_0\): female college graduates are more likely to report higher DOWN responses.  
Assumptions: independent observations, binomial outcome, and sufficiently large counts in each group.

```{r}
# (Optional) Show the same calculation in R
x <- c(6757, 11448)                 # successes (n_34)
n <- c(40663+6757, 60524+11448)     # totals
prop.test(x, n, alternative = "two.sided", correct = FALSE)


## We conducted two-proportion hypothesis tests to examine differences in higher DOWN responses (values 3–4). First, among college graduates, 14.25% of males and 15.90% of females reported higher DOWN values. The difference of 1.65 percentage points produced a z-statistic of about 6.1 with a p-value < 0.001, indicating a significant gender difference. Second, among females, 15.90% of college graduates and 26.6% of non–college graduates reported higher DOWN responses, with a z-statistic around 47 and p < 0.001, showing a strong education effect. Both tests assume random and independent samples, large sample sizes, and a binomially distributed outcome, supporting the validity of the results.