---
title: "Econometrics Exam 1"
author: "Romulo Jimenez"
date: "2025-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Overview

We compare the mean “feeling down” scores between adults with and without a college degree using a Welch two-sample t-test.  

### Summary Data

| Education | Mean | SD | N |
|------------|------|----|----|
| No College Degree | 1.987048 | 1.0543700 | 46942 |
| College Degree | 1.704870 | 0.9042484 | 120601 |

## 2. Hypothesis Test in R

```{r}
# Summary data
m1 <- 1.987048
s1 <- 1.0543700
n1 <- 46942

m2 <- 1.704870
s2 <- 0.9042484
n2 <- 120601

# Difference in means
diff <- m1 - m2

# Standard error (Welch)
SE <- sqrt(s1^2/n1 + s2^2/n2)

# Degrees of freedom (Welch-Satterthwaite)
df <- (s1^2/n1 + s2^2/n2)^2 / ((s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1))

# t-statistic, p-value, CI
t_stat <- diff / SE
p_val <- 2 * pt(-abs(t_stat), df = df)
CI <- diff + c(-1,1) * qt(0.975, df) * SE

list(
  difference = diff,
  SE = SE,
  df = df,
  t_stat = t_stat,
  p_value = p_val,
  CI_95 = CI
)

## The mean “feeling down” score for individuals without a college degree (1.99) was higher than for those with a college degree (1.70), producing an estimated difference of 0.28 points. The standard error of this difference was approximately 0.0056, yielding a t-statistic of about 50 and a p-value less than 0.001. Because the p-value is far below 0.05, we reject the null hypothesis that the two groups have equal means. We conclude that adults without a college degree report feeling “down” significantly more often than college graduates. The 95% confidence interval (approximately 0.27 to 0.29) confirms that the true mean difference is positive and statistically significant.







## Question 2

### Crosstab Summary

The table below summarizes the number of respondents by education level and gender:

| Education | Gender   | n_12 | n_34 |
|:----------:|:---------|:----:|:----:|
| no_coldeg  | male     | 13534 | 4425 |
| no_coldeg  | female   | 20840 | 7569 |
| no_coldeg  | tran/o   | 322 | 252 |
| coldeg     | male     | 40663 | 6757 |
| coldeg     | female   | 60524 | 11448 |
| coldeg     | tran/o   | 757 | 452 |

### Hypothesis Test

We conducted a two-proportion hypothesis test to examine whether the proportion of respondents reporting higher DOWN responses (values 3 or 4) differs between males and females with a college degree.  

The null hypothesis states that there is no difference in the proportion of higher DOWN responses between the two gender groups:

\[
H_0: p_{male} = p_{female}
\]

and the alternative hypothesis is:

\[
H_a: p_{male} \neq p_{female}
\]

Among college degree holders, **14.25% of males** and **15.90% of females** reported higher DOWN responses, yielding a difference of approximately **1.65 percentage points**.  
Using a pooled standard error of about 0.0027, the corresponding **z-statistic is ≈ 6.1**, with a **p-value < 0.001**.  

We therefore **reject the null hypothesis** and conclude that female respondents with a college degree are significantly more likely to report higher DOWN responses compared to males.  

For this test to be valid, we assume:
- Random and independent samples  
- Large enough counts in each group (at least 10 expected successes/failures)  
- A binomially distributed outcome (proportion data)
